---
title: "COVID Cases"
output: html_notebook
---
The following analysis was taking the US / Global Cases of COVID19 and using the data to draw some insights of the data. Specifically we are curious to see what the data tells us about 

```{r}
library(tidyverse)
library(lubridate)
```
#  following data comes from https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series


#https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series

```{r}
link <- "https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/"

files <- c("time_series_covid19_confirmed_US.csv", "time_series_covid19_confirmed_global.csv", "time_series_covid19_deaths_US.csv", "time_series_covid19_deaths_global.csv", "time_series_covid19_recovered_global.csv")

US_cases <- read_csv(paste0(link,files[1]))
global_cases <- read_csv(paste0(link,files[2]))
US_deaths <-  read_csv(paste0(link,files[3]))
global_deaths <- read_csv(paste0(link,files[4]))
# covid_data <- read_csv(link)

uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"
uid <- read_csv(uid_lookup_url)
```


# Cleaning Up the Data 
One of the most important parts of understanding the data is moldinging it into a fashion that would work for the users to read. 

## Cleaning up Global cases 
```{r}
global_cases  <- global_cases %>% pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), names_to = "date", values_to = "cases") %>% select(-c(Lat, Long))

global_deaths  <- global_deaths %>% pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), names_to = "date", values_to = "deaths") %>% select(-c(Lat, Long))
```


```{r}
global_deaths
```



## Joining data
```{r}
global <- global_cases %>% 
  full_join(global_deaths) %>% 
  rename(Country_Region = `Country/Region`, Province_State = `Province/State`) %>% 
  mutate(date = mdy(date))

global <-  global %>% filter(cases > 0)
```



```{r}
summary(global)
```

```{r}
global %>% filter(cases > 28000000)
```
```{r}

US_cases <- US_cases %>% pivot_longer(cols = -(UID:Combined_Key), 
                          names_to = "date", 
                          values_to = "cases") %>% 
  select(Admin2:cases) %>% mutate(date = mdy(date)) %>% 
  select(-c(Lat, Long_))
```

```{r}
US_deaths <- US_deaths %>% 
  pivot_longer(cols = -(UID:Population), 
               names_to = "date", 
               values_to = "deaths") %>% 
  select(Admin2:deaths) %>% 
  mutate(date = mdy(date)) %>% 
  select(-c(Lat, Long_))
```

joining together the two files to create a deaths file 
```{r}
US <- US_cases %>% full_join(US_deaths)
```

```{r}
global <- global %>% unite("Combined_Key", 
                           c(Province_State, Country_Region), 
                           sep = ", ", 
                           na.rm = TRUE, 
                           remove = FALSE)
```

Need to add population to the global dataset as well 

```{r}
uid <- uid %>% select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))
global <- global %>% left_join(uid, by = c("Province_State", "Country_Region")) %>% 
  select(-c(UID, FIPS)) %>% 
  select(Province_State, Country_Region, date, cases, deaths, Population, Combined_Key)

global 
```

```{r}
US_by_state <- US %>% 
  group_by(Province_State, Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  select(Province_State, Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% 
  ungroup()
```


```{r}
US_by_state
```


```{r}
US_totals <- US_by_state %>% 
  group_by(Country_Region, date ) %>% 
  summarize(cases = sum(cases), deaths = sum(deaths), 
            Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>% 
  select(Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% 
  ungroup()


```
```{r}
US_totals %>% filter(cases > 0) %>% 
  ggplot(aes(x = date, y = cases)) + 
  geom_line(aes(color = "cases")) + 
  geom_point(aes(color = "cases")) + 
  geom_line(aes(y = deaths, color = "deaths")) + 
  geom_point(aes(y = deaths, color = "deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(title = "COVID19 in US", y = NULL)
```
What about the states: 

```{r}

state <- "New York"
US_by_state %>% 
  filter(Province_State == state) %>%
  filter(cases > 0 ) %>% 
  ggplot(aes(x = date, y = cases)) +
  geom_line(aes(color="cases")) + 
  geom_point(aes(color="cases")) + 
  geom_line(aes(y =deaths, color = "deaths")) + 
  geom_point(aes(y = deaths, color = "deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) +
  labs(title=str_c("COVID 19 in ", state), y = NULL)
```

```{r}
max(US_totals$deaths)
```

```{r}
US_by_state <- US_by_state %>% 
  mutate(new_cases = cases - lag(cases))

US_totals <- US_totals %>%
  mutate(new_cases = cases - lag(cases), 
         new_deaths = deaths - lag(deaths))
```


